---
title: "Liam England research spring"
output: html_document
date: "2024-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
```


I am studying natural disasters. Specifically, I want to answer the question:
What is the relationship between displacement time and type of natural disaster for each state?
Identify the rate of natural distasters for each state generally as well

## Setting up

```{r, message=FALSE}
files<-list.files("data",recursive = TRUE,full.names = TRUE,pattern="puf")

pulse<-map_df(files,read_csv)
```

## Selecting the variables needed

```{r, message=FALSE}
# "                    '01'='Alabama'
#                     '02'='Alaska'
#                     '04'='Arizona'
#                     '05'='Arkansas'
#                     '06'='California'
#                     '08'='Colorado'
#                     '09'='Connecticut'
#                     '10'='Delaware'
#                     '11'='District of Columbia'
#                     '12'='Florida'
#                     '13'='Georgia'
#                     '15'='Hawaii'
#                     '16'='Idaho'
#                     '17'='Illinois'
#                     '18'='Indiana'
#                     '19'='Iowa'
#                     '20'='Kansas'
#                     '21'='Kentucky'
#                     '22'='Louisiana'
#                     '23'='Maine'
#                     '24'='Maryland'
#                     '25'='Massachusetts'
#                     '26'='Michigan'
#                     '27'='Minnesota'
#                     '28'='Mississippi'
#                     '29'='Missouri'
#                     '30'='Montana'
#                     '31'='Nebraska'
#                     '32'='Nevada'
#                     '33'='New Hampshire'
#                     '34'='New Jersey'
#                     '35'='New Mexico'
#                     '36'='New York'
#                     '37'='North Carolina'
#                     '38'='North Dakota'
#                     '39'='Ohio'
#                     '40'='Oklahoma'
#                     '41'='Oregon'
#                     '42'='Pennsylvania'
#                     '44'='Rhode Island'
#                     '45'='South Carolina'
#                     '46'='South Dakota'
#                     '47'='Tennessee'
#                     '48'='Texas'
#                     '49'='Utah'
#                     '50'='Vermont'
#                     '51'='Virginia'
#                     '53'='Washington'
#                     '54'='West Virginia'
#                     '55'='Wisconsin'
#                     '56'='Wyoming'"

disaster_data <- pulse %>%
  select(EST_ST, ND_DISPLACE:ND_HOWLONG)
```

## Generate basic statistics

```{r, message=FALSE}
disaster_data_bystate <- disaster_data %>%
  filter(ND_DISPLACE > 0) %>% # Only want respondents who answered the question
  group_by(EST_ST) %>%
  summarize(
    response_count = n(),
    displaced_by_disaster = sum(ND_DISPLACE == 1, na.rm=TRUE),
    percentage_displaced = displaced_by_disaster/response_count
  )

kable(head(disaster_data_bystate %>% arrange(desc(percentage_displaced))))
```

In 2023, the states that had the highest rate of displacement due to natural disaster were Louisiana (8%), Florida (5%), and Mississippi (1.8%).

Map of these results: 

```{r, message=FALSE}
library(tigris)
library(tmap)
library(sf)

us_geo <- tigris::states(class = "sf") 

disaster_data_bystate_map <- us_geo %>%
  left_join(disaster_data_bystate, by = c("STATEFP" = "EST_ST"))

tm_shape(disaster_data_bystate_map) +
  tm_polygons("percentage_displaced", style="jenks")
```


I ran out of time, but next I would have calculated and created maps of the most common type of natural disaster in each state.
